---
title: "Functionalizing NEON raster data"
author: "Sarah Graves"
date: "June 21, 2016"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Ojbectives

Start by putting up objectives/tasks that students will be working though:

1. **Import** a raster — A lidar canopy height model (lidar/Teak_lidarCHM.tif)
1. For the CHM, **set values** == 0 to NA (not trees)
1. **Visualize** density and plot vertical cutoff lines.
1. **Classify** the raster according to some distribution – low medium and tall trees. This could be done using a histogram potentially or we could just decide that <2m is generally grasses / understory, <6m small trees,and the rest are tall trees. A function could import the desired thresholds. Visualize histogram/density and plot vertical cutoff lines.
 (http://neon-workwithdata.github.io/neon-data-institute-2016/R/create-hillshade-R/)
1. **Plot** the classified raster, add a legend for each “class” - legends are super tricky to simplifying this process with a function would be good.  see: http://neon-workwithdata.github.io/neon-data-institute-2016/R/classify-by-threshold-R/  for my take on forcing a legend outside of the plot area using par settings. You may have other better forms of magic to make this work well. :)
1. **Export** the plot figure to a pdf – publishable
1. **Export** the classified raster as a geotiff with NaFlagg = -9999 to an outputs folder.

# Load packages

```{r LOAD-PACKAGES, message=FALSE}

rm(list = ls())

library(raster)

source("scripts/functions.R")

```


# Import raster

```{r IMPORT-CHM}

chm_file <- "../NEONdata/D17-California/TEAK/2013/lidar/TEAK_lidarCHM.tif"
chm <- raster(chm_file)
plot(chm,
     main="Raw canopy height model")

breaks <- c(5,10,20,100)
```


# Set NA values

```{r SET-NA}

par(mfrow=c(1,2))

hist(chm)

# assign NA to zeros because we don't want to analyze zero values (it's not vegetation)
chm[chm==0] <- NA

# check to see if zeros are removed
hist(chm)

par(mfrow=c(1,1))

```


# Set class breaks and classify raster

Classify the CHM to differentiate between short, medium, and tall vegetation.
Use a function to allow for any input of class breaks.

```{r CLASSIFY}

rcl.m <- create_ht_class_matrix(breaks)

```

```{r PLOT-DENS-WBREAKS}

plot_chm_dens(chm,title="CHM of TEAK with set breaks",breaks)

```

```{r PDF-SAVE-FUNCTION}

make_pdf(plot_chm_dens(chm,title="CHM of TEAK with set breaks",breaks),
         filename = "TEAK CHM density with breaks2.pdf",
         width=7,height=6)

```


# Classify raster

```{r CLASSIFY-RASTER}

# reclassify the raster using the reclass object - rcl.m
chm.class <- reclassify(chm,
                        rcl.m)

```


# Plot classified raster

```{r PLOT-RAST}

plot_reclassified_raster(chm,"TEAK",breaks)

```


# Export plot to PDF

```{r EXPORT-PDF}



```


# Export classified raster to GeoTiff

```{r EXPORT-TIF}

writeRaster(chm.class,"../outputs/TEAK/TEAK_chm_classified.tif",
            format="GTiff",
            NAflag=-9999,
            overwrite=T)

```


```{r SET-UP-INFO}

sessionInfo()

```

